Temporal Continuity and the Judgment of Actual Causation

### 2021-01-28 docs/resources/CogSci_paper/paper.tex
minor
added brms prior

To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms}. For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded. We then ran a logistic regression in which stativity (i.e.\ state vs event) and network length (long vs medium) were included as predictors of a binary dependent variable (root vs immediate). We included a random intercept for participant to allow for statistical dependencies between multiple responses from the same participant. The posterior mean of the intercept variable was 10.54 and the 95\% credible interval was $[2.30, 26.54]$, which supports the conclusion that root causes were chosen more often than immediate causes. The credible intervals for the other coefficients both included 0 (state: -2.75, $[-10.03, 1.14]$ ; long: 1.85, $[-1.78, 7.33]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.

@@

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E1}
%\includegraphics{figures/expt1.eps}
\includegraphics{figures/expt1test.eps}
\end{center}
\caption{Experiment 1. (a) Order of the four key events for dual branch sequences. The probability raising and continuity accounts both single out \ev{R}{C} as the main cause. (b)  Responses for dual branch sequences. The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence. Error bars show the standard error of the mean estimated using XXX bootstrap samples.} 
\label{fig:expt1}
\end{figure}

@@

\textbf{Participants}. 100 participants were recruited via Amazon Mechanical Turk and paid \$1 for a 6 minute experiment. Two did not submit a valid completion code, leaving 98 sets of responses for analysis.

@@

The second and more critical analysis used a binary dependent variable that indicated whether participants chose a cause on the continuous branch or the delayed branch.  The posterior mean of the intercept variable was 2.0 (95\% credible interval $[1.4, 2.65]$), indicating that the continuous branch was chosen more often than the delayed branch.  The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders. 
Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2. This difference may reflect the increased difficulty of Experiment 2. When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy of choosing the immediate cause that directly precedes the effect.

@@

The experimental paradigm we have developed can easily be extended to a more diverse set of causal systems. For example, by introducing preventive causal links, we can explore case of double prevention (e.g.\ cases where an active detector that would have prevented the activation of the GCM is itself inactivated by another detector). \citeA{glymourstone} point out that current work on actual causation is based largely on discussions of an ``infinitesimal fraction'' of the set of possible cases, but our particle detector task can support a comprehensive survey of people's judgments about actual causation.

>>>>

To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms} with default priors\footnote{By default brms uses improper flat priors on the coefficients of fixed effects, and a half student-t prior with 3 degrees of freedom, a location of 0 and a scale of 2.5 on the intercept and on the standard deviations of the random effects.}. For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded. We then ran a logistic regression in which stativity (i.e.\ state vs event) and network length (long vs medium) were included as predictors of a binary dependent variable (root vs immediate). We included a random intercept for participant to allow for statistical dependencies between multiple responses from the same participant. The posterior mean of the intercept variable was 10.54 and the 95\% credible interval was $[2.30, 26.54]$, which supports the conclusion that root causes were chosen more often than immediate causes. The credible intervals for the other coefficients both included 0 (state: -2.75, $[-10.03, 1.14]$ ; long: 1.85, $[-1.78, 7.33]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.

@@

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E1}
%\includegraphics{figures/expt1.eps}
\includegraphics{figures/expt1test.eps}
\end{center}
\caption{Experiment 1. (a) Order of the four key events for dual branch sequences. The probability raising and continuity accounts both single out \ev{R}{C} as the main cause. (b)  Responses for dual branch sequences. The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence. Error bars show the standard error of the mean estimated using 100 bootstrap samples.} 
\label{fig:expt1}
\end{figure}

@@

\textbf{Participants}. 100 participants were recruited via Amazon Mechanical Turk and paid \$1 for a 6 minute experiment. 
%Two did not submit a valid completion code, leaving 98 sets of responses for analysis. NO LONGER TRUE

@@

The second and more critical analysis used a binary dependent variable that indicated whether participants chose a cause on the continuous branch or the delayed branch.  The posterior mean of the intercept variable was 2.0 (95\% credible interval $[1.4, 2.65]$), indicating that the continuous branch was chosen more often than the delayed branch.  The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders. 
Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2. This difference may reflect the increased difficulty of Experiment 2. When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy of choosing the immediate cause that directly precedes the effect.

@@

The experimental paradigm we have developed can easily be extended to a more diverse set of causal systems. For example, by introducing preventive causal links, we can explore case of double prevention (e.g.\ cases where an active detector that would have prevented the activation of the GCM is itself inactivated by another detector). \citeA{glymourstone} point out that current work on actual causation is based largely on discussions of an ``infinitesimal fraction'' of the set of possible cases, but our paradigm can potentially be used to carry out a comprehensive survey of judgments about actual causation.
